---
title: "OSW_ggplot"
author: "T Haarman"
date: "5/20/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## OSW ggplot2

This R markdown documents contains examples of basic functions in ggplot2. In the workshop we will go through these examples to showcase the package. 

First we load the ggplot2 package.

```{r}
library(ggplot2)
```



## Simple scatterplot

Look at the data:

```{r}
mpg
```

Let's:
  1. create a basic scatterplot
  2. show the different car drives by colour/shape
  3. show the displacement values in the plot
  4. make the points a solid color
  5. add more geometries to the plot like a smoothed line or density layer

```{r}
ggplot(mpg, aes(x=cty, y=hwy)) +
  geom_point()

```

It matters where you put the mappings (and data)!

```{r}
ggplot(mpg, aes(x=hwy, y=cty, colour=drv)) +
  geom_point() +
  geom_smooth(method='lm')
```



## Let's look at some more geoms!

diamonds dataset:
```{r}
diamonds
```


Lets look at the distribution of a single variable (price)
```{r}
# Histogram
ggplot(diamonds, aes(x = price)) +
  geom_histogram()

# Area
# ggplot(diamonds, aes(x = price, fill=cut)) +
#   geom_area(stat="bin", bins=100)

# Density (kernel estimation)
# ggplot(diamonds, aes(x = price)) +
#   geom_density()

```




Now visualize two variables (carat vs price)
```{r}
ggplot(diamonds, aes(x = carat, 
                     y = price, 
                     colour = cut)) +
geom_point(alpha=0.1)  
# geom_rug() 
# geom_smooth()

```

With one categorical value: Box/violinplot
```{r}

ggplot(diamonds, aes(x = price, y = cut)) +
  geom_boxplot()
  # geom_violin()
```

Counts, with bar charts. We can fill it with the clarity to get a stacked bar chart.

```{r}
ggplot(diamonds, aes(x = cut)) +
  geom_bar()
```

But what if we already have the counts?
```{r}
#This is the same, but with precalculated counts
diamonds_counts <- diamonds %>% 
  group_by(cut) %>% 
  summarise(count = n())

ggplot(diamonds_counts, aes(x=cut, y=count)) +
  geom_bar(stat='identity')

# geom_col() is a shortcut for geom_bar(stat='identity')
ggplot(diamonds_counts, aes(x=cut, y=count)) +
  geom_col()

```

## Time series/line plots

You may need to do some data transformation on time series data to show what you want.
```{r}
eu_stocks <- data.frame(EuStockMarkets, date=time(EuStockMarkets))
eu_stocks_long <- eu_stocks %>% 
  pivot_longer(cols=c("DAX", "SMI", "CAC", "FTSE"))


ggplot(eu_stocks_long, aes(x = date, 
                           y = value,
                           color = name)) +
  geom_line()
```



## Scales
```{r}
ggplot(diamonds, aes(x = carat, 
                     y = price, 
                     colour = clarity)) +
  geom_point() +
  scale_colour_brewer(type='seq') +
  scale_x_continuous(trans="log10")

```


## Themes
Many predefined themes (e.g. bw/classic/dark/minimal). You can also define your own theme, for example on EOS you have ggDNB::theme_dnb().
```{r}
ggplot(mpg, aes(x = cty, 
                y = hwy, 
                colour = class)) +
  geom_point() +
  theme_classic()
```

```{r}
ggplot(mpg) + 
  geom_bar(aes(y = class, fill = drv)) +
  facet_grid(cols=vars(year)) +
  theme_minimal() +
  scale_fill_brewer(palette = 'Accent') +
  labs(title = 'Number of Cars by Year',
       caption = '© De Nederlandsche Bank',
       x = NULL,
       y = 'Car class',
       fill = 'Drive type') +
  theme(
    title = element_text(family = 'sans', 
                         face='italic', 
                         color='darkgray'),
    strip.text = element_text(face='bold', size=10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.x = element_line(linetype = 'dashed')
  )
```

```{r}
ggplot(mpg, aes(x=cty, y=displ)) +
  geom_point(aes(color=drv)) 
+
  geom_text(data=filter(mpg, hwy > 40), aes(label=model))
```






